Notes on symmetrizing the FDFD matrix with stretched-coordinate PML.

### 2012-02-04 scribblings/symmA.tex
First part of symmA done.

\documentclass{article}
\usepackage{amsmath}
\newcommand{\BE}{\begin{equation}}
\newcommand{\EE}{\end{equation}}
\newcommand{\D}{\Delta}
\begin{document}
\BE (AD)^T = DA^T \EE
\BE (AD)_{ij} = A_{ij} d_j \EE
\BE (DA^T)_{ij} = d_i (A^T)_{ij} = A_{ji} d_i \EE
Find $d$ such that $AD$ is symmetric where 
\BE A = \nabla \times \nabla \times \EE
for a non-regular cartesian grid.
In other words, find $d$ such that
\BE A_{ij} d_j = A_{ji} d_i, \quad \text{for all $i, j$.} \EE
\begin{quote} FIGURE HERE \end{quote}
For the figure above,
\BE A_{ba} = -\frac{1}{\D x_b \D y_b} \EE
\BE A_{ab} = -\frac{1}{\D x_a \D y_a} \EE
\BE d_i \propto \frac{1}{\D x_i \D y_i} \EE
Also, we will have
\BE A_{ca} = -\frac{1}{\D x_b \D z_b} \EE
\BE A_{ac} = -\frac{1}{\D x_a \D z_a} \EE
\BE d_i \propto \frac{1}{\D x_i \D z_i} \EE
Lastly, we already have $A_{aa'} = A_{a'a}$ 
    for $a$ and $a'$ in the same direction.

The conclusion, then, is
\BE d_i = \frac{1}{\D x_i \D y_i \D z_i} \EE

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{pstricks}
\newcommand{\BE}{\begin{equation}}
\newcommand{\EE}{\end{equation}}
\newcommand{\D}{\Delta}
\title{Notes on symmetrizing the FDFD matrix with stretched-coordinate PML.}
\author{Jesse Lu, Wonseok Shin}
\begin{document}
\maketitle
We want to find a diagonal matrix 
    \BE D = \text{diag}(d) \EE
    such that
    \BE (AD)^T = AD, \EE
    where $A = \nabla \times \nabla \times$ is the system matrix
    for the finite-difference frequency-domain problem
    for Maxwell's equations.

In fact, if we use an equally-spaced Cartesian grid 
    (constant $\D x$, $\D y$, and $\D z$),
    $A$ is already symmetric.
However, when applying a stretched-coordinate perfectly-matched layer (SC-PML)
    this no longer holds true,
    and one requires a diagonal preconditioning matrix $D$
    in order to restore the symmetry of $A$.
Note that we assume that both $A$ and $D$ are complex-valued,
    and that our goal is to make $A$ symmetric, not Hermitian.

First we note that,
    \BE (AD)^T = DA^T \EE
    \BE (AD)_{ij} = A_{ij} d_j \EE
    \BE (DA^T)_{ij} = d_i (A^T)_{ij} = A_{ji} d_i, \EE
    and so state our problem as finding $d$ such that
\BE A_{ij} d_j = A_{ji} d_i, \quad \text{for all $i, j$.} \EE

\begin{figure}[h]\begin{center}
\psset{gridcolor=green, subgridcolor=yellow}
\begin{pspicture}(5,5)
    \let\psgrid\relax
    \psline[linestyle=dotted](2,1)(4,1)
    \psline[linestyle=dotted](2,3)(4,3)
    \psline[linestyle=dotted](4,1)(4,3)
    \psline[linestyle=dotted](2,1)(2,3)
    \psline{->}(2.6,2)(3.4,2)
    \rput[t](3,1.9){$a$}
    \psline{|-|}(2,0.8)(4,0.8)
    \rput[t](3,0.7){$\D x_a$}
    \psline{|-|}(4.2,1)(4.2,3)
    \rput[l](4.3,2){$\D y_a$}

    \psline[linestyle=dotted](1,2)(1,4)
    \psline[linestyle=dotted](3,2)(3,4)
    \psline[linestyle=dotted](1,4)(3,4)
    \psline[linestyle=dotted](1,2)(3,2)
    \psline{->}(2,2.6)(2,3.4)
    \rput[r](1.9,3){$b$}
    \psline{|-|}(0.8,2)(0.8,4)
    \rput[r](0.7,3){$\D y_b$}
    \psline{|-|}(1,4.2)(3,4.2)
    \rput[b](2,4.3){$\D x_b$}
\end{pspicture}
\caption{Schematic used to find the values of $A_{ab}$ and $A_{ba}$,
            where $a$ is a vector component in the $x$-direction
            and $b$ is a vector component in the $y$-direction.
        A similar schematic can be used to determine
            the values for the $A_{ac}$ and $A_{ca}$,
            where $c$ is a vector component in the $z$-direction,
            by replacing $b \to c$ and $\D y_{a,b} \to \D z_{a,c}$.}
            \label{fig1}
\end{center} \end{figure}

To do so, we observe that from figure~\ref{fig1} we can deduce that the
    $\nabla \times \nabla \times$ operation will result in
    \BE A_{ba} = -\frac{1}{\D x_b \D y_b} \EE
    \BE A_{ab} = -\frac{1}{\D x_a \D y_a}, \EE
    from which we conclude that
    \BE d_i \propto \frac{1}{\D x_i \D y_i}. \EE

Similarly, if we replace $b \to c$ and $\D y_{a,b} \to \D z_{a,c}$ 
    in the figure, then we have
    \BE A_{ca} = -\frac{1}{\D x_b \D z_b} \EE
    \BE A_{ac} = -\frac{1}{\D x_a \D z_a}, \EE
    which leads to
    \BE d_i \propto \frac{1}{\D x_i \D z_i}. \EE

We then conclude that the choice of 
    \BE d_i = \frac{1}{\D x_i \D y_i \D z_i} \EE
    will lead to symmetric $AD$.
Note that none of the $A_{aa'}$ terms need to be included,
    where $a$ and $a'$ point in the same direction,
    since $A_{aa'} = A_{a'a}$ will hold both with or without the preconditioner.
